Verifying Hierarchical Cache Coherence Protocols: a theorem-prover based modular, reusable proof

### 2014-02-03 paper/System.tex
System Update for grammar/spelling

Each node maintains the a permission state for
each address, denoting what operations may be performed on the
data. An address may be in the modified state (\Mo) representing that
it may be both read and modified, or in the shared state (\Sh) where
it may only be read, or in the Invalid state (\In), meaning it can be
neither modified or read. This forms a natural total order of
permissions ($\In < \Sh < \Mo$). For leaf caches, the meaning of these
permission states is straightfoward, a load request can be processed only if the
leaf is at least in the \Sh{} state, and a store request can be processed only
if it is in the \Mo{} state. For the non-leaf cache, these states denote the
highest state present in the subtree rooted at this cache, denoting that some
cache in the subtree has the specified permissions (we are modeling inclusive
caches). Of course, all these invariants have to be enforced by the protocol.

@@

Non-leaf nodes also keep a copy of the coherence state of each of its
children in a \emph{directory}, which contains a the state of each of
its children caches for a particular address.  For a non-leaf cache
$p$, the state in its directory for child cache $c$ is denoted by
$p.\dstate[c][a]$, for address $a$. This local version may
be stale, in that the child may have changed it state but the parent
has not yet updated itself. However, the parent state will always be a
conservative overapproximation of the child state.

@@

The behavior of the system are described via atomic state transitions.
These transitions have two parts: the guard dictating when the
transition can fire, and the action defining how the state of the
system changes when the transition happens. These transitions or \emph{guarded
atomic actions} logically happen one by one, making the overall behavior of the
system well defined. Figure~\ref{trans} gives the atomic transitions for a
cache coherent memory subsystem. The actions in the transitions repesent only
the specific system states that change. We use $x \Leftarrow y$ to indicate
setting of a specific state $x$ of the overall system to value $y$ because of a
transition.

Figure~\ref{childside} shows the transitions that result in an upgrade of the
state of a cache. It first involves a cache sending an upgrade request to the
its parent when the cache is not already waiting for a response for that
address from the parent (ChildSendReq). This request is received by the parent
(ParentRecvReq), and if the rest of the parent's children caches are
\emph{compatible}, and the parents state also has the permission being
requested, it sends an upgrade response to the cache, while changing the
appropriate directory state. Compatibility of an upgrade request with respect
to a cache's siblings can be defined as follows:
\begin{multline*}
\compat = \lambda (p \in Cache) (c \in Cache) (a \in \textit{Addr}) (s \in
\{\Mo, \Sh, \In\}) \Rightarrow\\ \forall i, i \ne c @->
\mylet s_i := p.\dstate[i][a] \myin\\
 (s = \Mo @-> s_i = \In) \wedge (s = \Sh @-> s_i = \Sh) \wedge (s = \In @-> s_i = \Mo)
\end{multline*}
Finally, the child cache receives the response from the parent (ChildRecvResp)
and sets its state while reseting its wait state if the desired upgrade is
indicated by the response.

Figure~\ref{parentside} shows the transitions that result in a downgrade of the
state of a cache because of a request by the cache's parent when the parent is
not already waiting for a response from that cache (ParentSendReq).  This
request is received by the child (ChildRecvReq), and if the state of each of
the child's own children is not greater than the requested downgrade, the cache
downgrades its state and sends a notification response to the parent. Finally,
the parent receives the response from the cache (ParentRecvResp) and sets the
appropriate directory state while reseting the appropriate directory wait state
if the desired downgrade is indicated by the response.

@@

%%%%%%%%%%%%%%%%%%%%%%%

%In addition, the leaf caches (called L1 caches usually)
%communicate with the processor through two channels: a) \cproc{} \textbf{FIFO}
%channel to transmit requests from the processor to L1 cache, and b) \procc{}
%\textbf{FIFO} channel to transmit responses from the L1 cache to the processor.
%There is exactly one of \cproc{} and \procc{} channels between the L1 and the processor.

>>>>

Each node maintains the a permission state for
each address, denoting what operations may be performed on the
data. An address may be in the modified state (\Mo) representing that
it may be both read and modified, or in the shared state (\Sh) where
it may only be read, or in the Invalid state (\In), meaning it can be
neither modified or read. This forms a natural total order of
permissions ($\In < \Sh < \Mo$). For leaf caches, the meaning of these
permission states is straightforward, a load request can be processed only if the
leaf is at least in the \Sh{} state, and a store request can be processed only
if it is in the \Mo{} state. For the non-leaf cache, these states denote the
highest state present in the subtree rooted at this cache, denoting that some
cache in the subtree has the specified permissions (we are modeling inclusive
caches). Of course, all these invariants have to be enforced by the protocol.

@@

Non-leaf nodes also keep a copy of the coherence state of each of its
children in a \emph{directory}, which contains a the state of each of
its children caches for a particular address. For a non-leaf cache
$p$, the state in its directory for child cache $c$ is denoted by
$p.\dstate[c][a]$, for address $a$. This local version may be stale,
in that the child may have changed it state but the parent has not yet
updated itself. However, the parent state will always be a
conservative over-approximation of the child state.

@@

The behavior of the system are described via atomic state transitions.
These transitions have two parts: the guard dictating when the
transition can fire, and the action defining how the state of the
system changes when the transition happens. These transitions or \emph{guarded
atomic actions} logically happen one by one, making the overall behavior of the
system well defined. Figure~\ref{trans} gives the atomic transitions for a
cache coherent memory subsystem. The actions in the transitions represent only
the specific system states that change. We use $x \Leftarrow y$ to indicate
setting of a specific state $x$ of the overall system to value $y$ because of a
transition.

Figure~\ref{childside} shows the transitions that result in an upgrade of the
state of a cache. It first involves a cache sending an upgrade request to the
its parent when the cache is not already waiting for a response for that
address from the parent (ChildSendReq). This request is received by the parent
(ParentRecvReq), and if the rest of the parent's children caches are
\emph{compatible}, and the parents state also has the permission being
requested, it sends an upgrade response to the cache, while changing the
appropriate directory state. Compatibility of an upgrade request with respect
to a cache's siblings can be defined as follows:
\begin{multline*}
\compat = \lambda (p \in Cache) (c \in Cache) (a \in \textit{Addr}) (s \in
\{\Mo, \Sh, \In\}) \Rightarrow\\ \forall i, i \ne c @->
\mylet s_i := p.\dstate[i][a] \myin\\
 (s = \Mo @-> s_i = \In) \wedge (s = \Sh @-> s_i = \Sh) \wedge (s = \In @-> s_i = \Mo)
\end{multline*}
Finally, the child cache receives the response from the parent (ChildRecvResp)
and sets its state while resetting its wait state if the desired upgrade is
indicated by the response.

Figure~\ref{parentside} shows the transitions that result in a downgrade of the
state of a cache because of a request by the cache's parent when the parent is
not already waiting for a response from that cache (ParentSendReq).  This
request is received by the child (ChildRecvReq), and if the state of each of
the child's own children is not greater than the requested downgrade, the cache
downgrades its state and sends a notification response to the parent. Finally,
the parent receives the response from the cache (ParentRecvResp) and sets the
appropriate directory state while resetting the appropriate directory wait state
if the desired downgrade is indicated by the response.

@@
